---
title: "covid"
author: "Chenghui Zheng"
date: "11/11/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setlocale("LC_ALL","English")
```

```{r lib, message=FALSE, warning=FALSE}
# Load all the libraries needed.
library(lubridate)
library(tidyverse) # to tidy data
library(dplyr) #data manipulation
library(gridExtra) # used to arrange plot
library(ggplot2)
library(zoo)
library(dplyr)
library(tidyr)
library(data.table)
library(forecast)
```

## Data Import

```{r data.import, message=FALSE}
# policy data from link cihi
# ref: https://www.cihi.ca/en/covid-19-intervention-timeline-in-canada#resources
policy <- readr::read_csv("policy_sept.csv")
# case data from ontario covid data
# ref: https://covid-19.ontario.ca/data
cases <- readr::read_csv("cases.csv") %>%
  mutate(date=lubridate::mdy(sub("^(\\w+)\\s?(.*)$","\\2",category))) %>%
  setnames(old = c('New cases','7-day average', 'Total cases'), 
           new = c('new_cases','avg_cases', 'total_cases')) %>%
  subset(select = c(date,new_cases,avg_cases, total_cases))
death <- readr::read_csv("death.csv") %>%
  mutate(date=lubridate::mdy(sub("^(\\w+)\\s?(.*)$","\\2",category))) %>%
  setnames(old = c('New deaths'), new = c('death')) %>%
  subset(select = c(date,death))
hospital <- readr::read_csv("activeHospital.csv") %>%
  mutate(date=lubridate::mdy(sub("^(\\w+)\\s?(.*)$","\\2",category))) %>%
  setnames(new = c('category','active_cases','hospital_cases','date')) %>%
  subset(select = c(date,active_cases,hospital_cases))
icu <- readr::read_csv("Hospital.csv") %>%
  mutate(date=lubridate::mdy(sub("^(\\w+)\\s?(.*)$","\\2",category))) %>%
  setnames(new = c('category','hospital_cases','icu_cases','date')) %>%
  subset(select = c(date,icu_cases)) # hospital is duplicate
test <- readr::read_csv("testsDaily.csv") %>%
  mutate(date=lubridate::mdy(sub("^(\\w+)\\s?(.*)$","\\2",category))) %>%
  setnames(new = c('category','tests','pos_perc','date')) %>%
  subset(select = c(date,tests,pos_perc))
# join all to case_on
case_on <- cases %>%
  left_join(death, by="date") %>%
  left_join(hospital, by="date") %>%
  left_join(icu,by="date") %>%
  left_join(test,by="date")
```

```{r ts_case}
case_ts <- zoo(case_on$new_cases, seq(from = as.Date("2020-01-30"), to = as.Date("2020-08-10"), by = 1)) # change end date if needed
case <- case_on$new_cases
# policies of Canada and Ontario
policy_on <- policy %>%
  filter(Jurisdiction == "Ont." | Jurisdiction == "Can.") %>% 
  mutate(Province = ifelse(Jurisdiction == "Ont.","Ontario","Canada")) %>%
  # prepare date field
  mutate(date = as.Date(ifelse(`Date implemented`=="No data" & 
                                 `Date announced`=="No data", NA,
                               ifelse(`Date implemented`=="No data", 
                                      `Date announced`, 
                                      `Date implemented`))))
```


```{r}
policy_on <- policy_on %>% setnames(old = "Intervention category", new = "intervention_category")
ggplot(policy_on, aes(x=intervention_category)) +geom_bar()+theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```



```{r}
policy_on <- policy_on %>% setnames(old = "Date implemented", new = "dat_implemented")
day_imple_policy <- tibble(policy_on %>% group_by(`dat_implemented`) %>% summarise(count = n())) %>% mutate(dat_implemented = as.Date(dat_implemented))

on_policy_case <-day_imple_policy %>% left_join(case_on, by = c("dat_implemented" ="date"))

new_on_policy_case <- on_policy_case %>% left_join(policy_on, by=c("dat_implemented" = "date"))
```


```{r}
#infection rate
new_on_policy_case1 <- new_on_policy_case %>% drop_na(new_cases) %>% filter(new_cases >0)
new_on_policy_case2 <- new_on_policy_case %>% drop_na(total_cases)

library(MASS)
boxcox(total_cases~ intervention_category, data=new_on_policy_case2)

lm.1 <- lm(log(total_cases)~ intervention_category, data=new_on_policy_case2)
summary(lm.1)


#boxcox(new_cases~ intervention_category, data=new_on_policy_case1)
lm.2 <- lm(sqrt(new_cases)~ intervention_category, data=new_on_policy_case1)
summary(lm.2)

lm.3 <- lm(log(new_cases)~ intervention_category, data=new_on_policy_case1)
summary(lm.3)

```


```{r}
ggplot(new_on_policy_case, aes(intervention_category, dat_implemented)) + geom_violin()+theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```


```{r}
ggplot(new_on_policy_case,aes(y=new_cases,x=intervention_category))+geom_boxplot()+theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

```



```{r}
library(lavaan)
library(semPlot)
library(OpenMx)
library(tidyverse)
library(knitr)
library(kableExtra)
library(GGally)
```
#path analysis: path analysis is used to describe the directed dependencies among a set of variables.
#A path coefficient is a standardized regression coefficient
```{r}
library(lavaan)
library(semPlot)
library(OpenMx)
library(knitr)
library(kableExtra)
library(GGally)
#-0.14
#unique(new_on_policy_case$intervention_category)
new_on_policy_case <- new_on_policy_case %>% filter(new_cases >0) %>% mutate(log_new_case = log(new_cases))
reduced_on_policy_case <- new_on_policy_case %>% filter(intervention_category %in% c("Travel", "State of emergency", "Closures", "Distancing", "Financial and economic"))
model <- 'log_new_case ~ intervention_category'
fit_reduce <- cfa(model, data=reduced_on_policy_case)
fit_all <- cfa(model, data=new_on_policy_case)
#summary(fit, fit.measures = TRUE, standardized = TRUE, rsquare = TRUE)

semPlot::semPaths(fit_reduce, "std",layout = 'circle',
             sizeMan = 15, sizeInt = 15, sizeLat = 15,
             edge.label.cex=1.5,
             fade=FALSE)

semPlot::semPaths(fit_all, "std",layout = 'circle',
             sizeMan = 15, sizeInt = 15, sizeLat = 15,
             edge.label.cex=1.5,
             fade=FALSE)


```


```{r}


on_policy_case <- on_policy_case %>% slice(16:40)

 ggplot(on_policy_case, aes(count, new_cases)) +
     geom_path(colour = "grey50") +
     geom_point(aes(colour = dat_implemented))
```



```{r plot}
autoplot.zoo(case_ts) + 
  geom_vline(xintercept = as.Date('2020-03-17'), color="red") +
  geom_vline(xintercept = as.Date('2020-05-19'), color="blue") +
  geom_vline(xintercept = as.Date('2020-06-12'), color="blue") +
  geom_vline(xintercept = as.Date('2020-07-17'), color="blue") +
  ggtitle("Ontario COVID-19 New Cases") + ylab("Cases") + xlab("Date") +
  geom_text(aes(x=as.Date('2020-03-17'), y=600), label="Ontario Entered Emergency State", angle=90, text=element_text(size=11)) +
  geom_text(aes(x=as.Date('2020-05-19'), y=600), label="Stage 1 Re-opening", angle=90, text=element_text(size=11)) +
  geom_text(aes(x=as.Date('2020-06-12'), y=600), label="Stage 2 Re-opening", angle=90, text=element_text(size=11)) +
  geom_text(aes(x=as.Date('2020-07-17'), y=600), label="Stage 3 Re-opening", angle=90, text=element_text(size=11))
```

```{r acf}
# find parameters for ARIMA model
acf(case)
diff1=diff(case,differences=1,lag=1)
acf(diff1)
pacf(diff1)
```

After doing first order differencing, it could be noticed that the acf already drops to a low value after lag 1 and could be viewed as exponential decay. And the partial acf cuts off after lag 3. Hence, it would be appropriate to propose an ARIMA model of ARIMA(p=3, d=1, q=0).

```{r arima1}
# first split the data before/after policy
#arima1 = sarima(logtrain,p=3,d=2,q=0)
```


```{r dataprep}
policy_cnt_type <- policy_on %>%
  dplyr::count(date,`Intervention category`) %>% 
  tidyr::spread(key =`Intervention category`,value = n)
policy_cnt_level <- policy_on %>%
  dplyr::count(date,Province) %>% 
  tidyr::spread(key = Province,value = n)
# perform join with case table
df <- case_on %>%
  left_join(policy_cnt_level, by ='date') %>%
  left_join(policy_cnt_type, by = 'date') 
# fill in all missing values with 0
df[is.na(df)] <- 0
# add cumulative column for each type and jurisdiction
df <- df %>%
  within(acc_can <- cumsum(Canada)) %>%
  within(acc_on <- cumsum(Ontario)) %>%
  within(acc_casemgt <- cumsum(`Case management`)) %>%
  within(acc_clo <- cumsum(Closures)) %>%
  within(acc_cinfo <- cumsum(`Contextual information`)) %>%
  within(acc_dist <- cumsum(Distancing)) %>%
  within(acc_fin <- cumsum(`Financial and economic`)) %>%
  within(acc_hserv <- cumsum(`Health services`)) %>%
  within(acc_hwork <- cumsum(`Health workforce`)) %>%
  within(acc_open <- cumsum(Openings)) %>%
  within(acc_pubinfo <- cumsum(`Public information`)) %>%
  within(acc_soe <- cumsum(`State of emergency`)) %>%
  within(acc_trav <- cumsum(Travel))
df_train <- df %>%
  filter(date <= as.Date("2020-08-10")) # policy data only available till Aug 10
df_test <- df %>%
  filter(date > as.Date("2020-08-10"))
```


```{r regARIMA}
# ARIMA model with external regressors
Xreg = data.matrix(df_train[c(25:35)])
fit <- auto.arima(df_train$new_cases, xreg = Xreg, stepwise = TRUE, 
                  seasonal = FALSE)
```


```{r arima}
summary(fit)
```

```{r pval}
(1-pnorm(abs(fit$coef)/sqrt(diag(fit$var.coef))))*2
```

```{r plot}
plot(seq(as.Date("2020/1/30"), by = "day", length.out = 184),
     fit$fitted, type = 'l')
lines(seq(as.Date("2020/1/30"), by = "day", length.out = 184),
      df_train$new_cases, col="red")
```